### 2013-02-14 doc/xps2013/proposal.tex
for the win

\paragraph{Hardware support for aggregation:} If the increased latency caused
by extra hops in hierarchical aggregation turns out to be detrimental, we
will explore mitigation by moving some of the aggregation logic to the network
card, either by exploiting off-the-shelf programmable network cards or
experimental systems such ask NetFPGA~\cite{netfpga} (while we are planning to
focus on Infiniband-based networking, if we explore the NetFPGA route we will
be using Ethernet).

@@

This interface suffers from multiple shortcomings. The biggest of which is
that ultimately, moving data to and from a remote node constructively requires
higher level semantic operations than simple reads and writes. These higher
level semantics can, of course, be synthesized from the underlying read and
write primitives, but it wastes round trips and synchronization time. One
possible alternative solution is to enable remote nodes to dispatch high-level
data-structure manipulation requests (e.g. insert this item into a queue,
remove an item from a set). The challenge here is to make those calls
programmable and interoperate securely with a modern operating system, and a
user-level runtime system. Additionally there are questions about \emph{who}
should carry out the computation. Classic RDMA does not interrupt the host
processor, but these richer programmable semantics will require either a
processor in the network interface (necessitating issues such as coherence be
managed) or complex RDMA tasks need a mechanism to interrupt the host
processor \emph{without leaving user space}.

@@

% In the long run, the solution here likely
% lies at the intersection of what progress is occurring in processor support
% for virtual machines and virtual memory translation for I/O. Our goal will
% be to study carefully the I/O virtualization proposals put forth by major
% processor manufactures and devise solutions that enable RDMA to capitalize
% on these developments.

@@

%% Sorry Mark, while the item below is clearly rich and juicy, we can't leave it bhere because it is huuuuge, it does not match the scope of the grant, and we dont have the expertive --- e.g., are you sure nobody is working on it?

%% I think this is the only idea that is going to work... maybe we should delete the above three and make this one good.  Seriously, the ideas up there are bunk.  -ark
% \paragraph{Rearchitecting flash:} Clearly the best solution here lays in re-architecting the flash chip itself.  We will explore new flash microarchitectures that optimize for message rate as well as bandwidth.  Such chips have value far beyond Grappa.  Research into persistent key/value stores, and really any persistent high-throughput data-structure will benefit from a redesigned flash chip that can sustain dramatically increased message rates.

@@

\paragraph{Task-weaving: } Much of the time the same task code is executing for hundreds of thousands of tasks in the Grappa runtime system.  For example, each task may be processing a loop body for a parallel for loop.  However, tasks introduce overhead, so we already manually impose a cutoff on recursive decomposition of parallel loops.  This way, multiple loop bodies are collected into a single task and executed serially.  However, this limits the amount of parallelism available for latency tolerance. A mmore efficient technique is to allow the compiler to unroll and schedule all loop bodies simultaneously within the same task, including overlapping remote memory operations, thereby ``weaving'' these independent tasks together to overlap computation with communication with minimal task overhead.  Further, a set of ``woven tasks'' may be amenable to vector execution, allowing even more parallelism to be exploited.

\paragraph{Task load control:} Having too few tasks per core fails to tolerate
fully the latency; having too many increases overheads unnecessarily.
Currently, we choose a number at program instantiation: experiments so far
have shown evidence of a broad sweet-spot, though it is of course varies as a
function of application behavior and input data. At compile-time, we might be
able to estimate the number of tasks required based on the density of remote
operations and system parameters (e.g., memory size, network latency, etc.);
at runtime we can adjust the number of tasks dynamically based on periodic
performance assessments.

% Presently we are starting work on this language and compiler support.  We are basing our effort on the LLVM~\cite{llvm} infrastructure.

@@

This benchmark suite will consist of application kernels, reference implementations, data-sets, data-set generators, documentation, specifications for generating standardized performance results, and a hosting service (http://graphbench.org) to provide public access to the benchmark and datasets as well as aggregate and publish results by all users of the benchmark suite.  Specifically, we aim to provide transparency and an opportunity to build off of others' discoveries by requiring source code be provided with any submission.

@@

We will build a collection of key computational kernels from graph and irregular applications. The benchmark suite will consist of both documentation on the actual kernel (ideally a research paper or book reference sourced from the community), and a reference implementation. The benchmark suite will also include sample data collected from existing sources, and synthetic workload generators. The collected data provides a grounding for performance analysis on real-world problems. The workload generators provide cleaner, easier to reason about randomly generated data-sets at a programmable scale.

>>>>

\paragraph{Hardware support for aggregation:} If the increased latency caused by extra hops in hierarchical aggregation turns out to be detrimental, we will explore moving some of the aggregation logic to the network card, either by exploiting off-the-shelf programmable network cards or experimental systems such as NetFPGA~\cite{netfpga} (while we are planning to focus on infiniband-based networking, if we use the NetFPGA route we will be using Ethernet as a test-bed).

@@

This interface suffers from multiple shortcomings. The biggest of which is that ultimately, moving data to and from a remote node constructively requires higher level semantic operations than simple reads and writes. These higher level semantics can, of course, be synthesized from the underlying read and write primitives, but it wastes round trips and synchronization time. One possible alternative solution is to enable remote nodes to dispatch high-level data-structure manipulation requests (e.g. insert this item into a queue, remove an item from a set). Past work on Active Messages~\cite{vonEicken92} provides the foundation on which to build these primitives.  The challenge here is to make those calls programmable and interoperate securely with a modern operating system, and a user-level runtime system. Additionally there are questions about \emph{who} should carry out the computation. Classic RDMA does not interrupt the host processor, but these richer programmable semantics will require either a processor in the network interface or complex RDMA tasks need a mechanism to interrupt the host processor \emph{without leaving user space}.

@@

@@

@@

\paragraph{Task-weaving: } Much of the time the same task code is executing for hundreds of thousands of tasks in the Grappa runtime system.  For example, each task may be processing a loop body for a parallel for loop.  However, tasks introduce overhead, so we already manually impose a cutoff on recursive decomposition of parallel loops.  This way, multiple loop bodies are collected into a single task and executed serially.  However, this limits the amount of parallelism available for latency tolerance. A more efficient technique is to allow the compiler to unroll and schedule all loop bodies simultaneously within the same task, including overlapping remote memory operations, thereby ``weaving'' these independent tasks together to overlap computation with communication with minimal task overhead.  Further, a set of ``woven tasks'' may be amenable to vector execution, allowing even more parallelism to be exploited.

\paragraph{Task load control:} Having too few tasks per core fails to fully tolerate remote memory latency; having too many increases overheads unnecessarily. Currently, we choose a fixed number at program instantiation.  This approach is sufficient as a first step, but isn't a workable long term option.  Our goal will be to develop as part of the runtime, auto-tuning technology that dynamically controls the number of spawned tasks.

@@

This benchmark suite will consist of application kernels, reference implementations, data-sets, data-set generators, documentation, specifications for generating standardized performance results, and a hosting service (http://graphbench.org) to provide public access to the benchmark and datasets as well as aggregate and publish results by all users of the benchmark suite.  We will start by adapting and building a collection of key computational kernels from graph and irregular applications.  We will collect sample data from existing sources, and produce synthetic workload generators. The collected data provides a grounding for performance analysis on real-world problems. The workload generators provide cleaner, easier to reason about randomly generated data-sets at a programmable scale.

@@
